---
title: "Essential R Programming"
author: 郭耀仁 Yao-Jen Kuo
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    center: true
    reveal_options:
      self_contained: true
      previewLinks: true
      slideNumber: true
---

# Basic Syntax

## The symbol of assignment `<-`

```{r}
hello_msg <- "Hello from R!"
hello_msg
```

## The symbol of comments `#`

```{r}
# Saying hello from R
hello_msg <- "Hello from R!"
hello_msg # print hello_msg to console
```

## Style guide for R programming

<http://adv-r.had.co.nz/Style.html>

## Useful built-in functions

- `print()`
- `help()`
- `class()`
- `rm()`
- `getwd()`
- `sessionInfo()`

## Common vector types

- numeric
- character
- logical
- Date

## Numeric with length of 1

```{r}
numeric_vector <- 5566
numeric_vector
class(numeric_vector)
length(numeric_vector)
```

## Numeric with length larger than 1

```{r}
numeric_vector <- c(5, 5, 6, 6) # using help(c) or ?c to check what c() function is
numeric_vector
class(numeric_vector)
length(numeric_vector)
```

## Character with length of 1

```{r}
character_vector <- "五五六六"
character_vector
class(character_vector)
length(character_vector)
```

## Character with length larger than 1

```{r}
character_vector <- c("五", "五", "六", "六")
character_vector
class(character_vector)
length(character_vector)
```

## Logical with length of 1

```{r}
logical_vector <- TRUE
logical_vector
class(logical_vector)
length(logical_vector)
```

## Date with length of 1

```{r}
date_vector <- Sys.Date()
date_vector
class(date_vector)
length(date_vector)
```

## Date with length larger than 1

```{r}
date_vector <- c(Sys.Date(), Sys.Date() - 1, Sys.Date() - 2, Sys.Date() - 3)
date_vector
class(date_vector)
length(date_vector)
```

## if statement

```{r eval=FALSE}
if (LOGICAL) {
  # Codes written here will be ran once LOGICAL is evaluated as TRUE
  # ...
  # ...
}
```

---

```{r}
wanna_fight <- TRUE
if (wanna_fight) {
  print("要打，就去練舞室打！")
}
```

## if-else statement

```{r eval=FALSE}
if (LOGICAL) {
  # Codes written here will be ran once LOGICAL is evaluated as TRUE
} else {
  # Codes written here will be ran once LOGICAL is evaluated as FALSE
}
```

---

```{r}
wanna_fight <- FALSE
if (wanna_fight) {
  print("要打，就去練舞室打！")
} else {
  print("那就不要打架。")
}
```

## if-else if-else statement

```{r eval=FALSE}
if (LOGICAL_1) {
  # Codes written here will be ran once LOGICAL_1 is evaluated as TRUE
} else if (LOGICAL_2) {
  # Codes written here will be ran once LOGICAL_1 is evaluated as FALSE
  # And LOGICAL_2 is evaluated as TRUE
} else {
  # Codes written here will be ran once both LOGICAL_1 and LOGICAL_2 are evaluated as TRUE
}
```

---

```{r}
wanna_fight <- FALSE
dancing_room_is_open <- FALSE
if (wanna_fight) {
  print("要打，先看練舞室有沒有開放。")
} else if (dancing_room_is_open) {
  print("要打，練舞室又有開放就去練舞室打！")
} else {
  print("那就不要打架。")
}
```

# Built-in Data Structures

## An overview of built-in data structures

|Number of Dimensions|Homogeneous|Heterogeneous|
|--------------------|-----------|-------------|
|1|Vector|List|
|2|Matrix|Data frame|
|n|Array||

## Using functions to create various data structures

## Using `list()` to create lists

```{r}
my_list <- list(
  c(5, 5, 6, 6),
  c(TRUE, TRUE, FALSE, FALSE),
  "五五六六"
)
my_list
class(my_list)
```

## Using `matrix()` to create matrices

```{r}
my_matrix <- matrix(c(1, 2, 3, 4, 5, 6), nrow = 3, ncol = 2)
my_matrix
class(my_matrix)
```

## Using `data.frame()` to create data frames

```{r}
my_df <- data.frame(
  title = c("The Shawshank Redemption", "The Dark Knight", "Forrest Gump", "Inception"),
  year = c(1994, 2008, 1994, 2010),
  imdb_rating = c(9.2, 9, 8.8, 8.7)
)
my_df
class(my_df)
```

## Using `array()` to create arrays

```{r}
my_array <- array(1:24, dim = c(4, 3, 2))
my_array
class(my_array)
```

# Intermediate Syntax

## Using `[]` to index a vector

```{r}
numeric_vector <- 6:10
numeric_vector[1]
numeric_vector[2]
numeric_vector[3]
numeric_vector[4]
numeric_vector[5]
```

## Using `[[]]` to index a list

```{r}
my_list <- list(
  title = c("The Shawshank Redemption", "The Dark Knight", "Forrest Gump", "Inception"),
  year = c(1994, 2008, 1994, 2010),
  imdb_rating = c(9.2, 9, 8.8, 8.7)
)
my_list[[1]] # my_list[["title"]]
my_list[[2]] # my_list[["year"]]
my_list[[3]] # my_list[["imdb_rating"]]
```

## Iterating with `for`

```{r eval=FALSE}
for (i in ITERABLE) {
  # Codes written here will be ran repeatedly
  # until i reaches the end of ITERABLE
}
```

## Iterating over a vector with `for`

```{r}
numeric_vector <- 6:10
for (i in numeric_vector) {
  print(i)
}
```

## Iterating over a list with `for`

```{r}
my_list <- list(
  title = c("The Shawshank Redemption", "The Dark Knight", "Forrest Gump", "Inception"),
  year = c(1994, 2008, 1994, 2010),
  imdb_rating = c(9.2, 9, 8.8, 8.7)
)
for (i in my_list) {
  print(i)
}
```

## Iterating with `while`

```{r eval=FALSE}
i <- 1
while (LOGICAL) {
  # Codes written here will be ran repeatedly
  # until i reaches the end of ITERABLE
  i <- i + 1
}
```

## Iterating over a vector with `while`

```{r}
numeric_vector <- 6:10
i <- 1
while (i <= length(numeric_vector)) {
  print(numeric_vector[i])
  i <- i + 1
}
```

## Iterating over a list with `while`

```{r}
my_list <- list(
  title = c("The Shawshank Redemption", "The Dark Knight", "Forrest Gump", "Inception"),
  year = c(1994, 2008, 1994, 2010),
  imdb_rating = c(9.2, 9, 8.8, 8.7)
)
i <- 1
while (i <= length(my_list)) {
  print(my_list[[i]])
  i <- i + 1
}
```

## Defining functions

```{r eval=FALSE}
function_name <- function(INPUT_1, INPUT_2, INPUT_3, ...) {
  # Codes written here can be ran once the function is called
  # ...
  # ...
  return(OUTPUT)
}
```

## Defining a function to convert celsius to fahrenheit

```{r}
celsius_to_fahrenheit <- function(degree) {
  degree_f <- degree * 9/5 + 32
  return(degree_f)
}
celsius_to_fahrenheit(30)
```

## Applying a function over a vector

```{r}
celsius_degrees <- 30:35
celsius_to_fahrenheit(celsius_degrees)
```

## Applying a function over a list

```{r}
celsius_degrees <- list(30, 31, 32, 33, 34, 35)
lapply(celsius_degrees, FUN=celsius_to_fahrenheit) # lapply as in list apply
```