---
title: "Data in Statistics"
author: 郭耀仁 Yao-Jen Kuo
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    center: true
    reveal_options:
      self_contained: true
      previewLinks: true
      slideNumber: true
---

# Univariate data

## What is univariate data

A univariate data set is a set of measurements for some variable from a collection of cases.

$$
x_1, x_2, ..., x_n
$$

## How do we use R for univariate data

Vector.

## Getting a vector with real-world data

```{r message=FALSE}
library(quantmod)

start_date <- "2018-01-01"
getSymbols(Symbols = "VTI", from = start_date) # create a VTI xts object in global environment
vti <- as.vector(coredata(VTI$VTI.Adjusted)) # extract close prices as a vector
class(vti)
```

## Summaries of a numeric vector

- Center
- Spread
- Shape

## Center: using `mean()` for sample mean

$$
sample \: mean = \bar{x} = \frac{1}{n}(x_1 + x_2 + ... + x_n) = \frac{1}{n}\sum_i x_i
$$

```{r}
mean(vti)
```

## Center: using `median()` for sample median

The middle value of the sorted vector.

```{r}
median(vti)
```

## Center: using `quantile()` for a generalized version of `median()`

```{r}
quantile(vti)
```

## Spread: using `range()` for the difference between the smallest and largest values

```{r}
range(vti)
diff(range(vti))
```

## Spread: using `IQR` for interquantile range

While range is an obvious measure of spread, but that it suffers from being effected by just one large or small value.

```{r}
IQR(vti)
```

## Spread: using `var()` for sample variance

$$
sample \; variance = s^2 = \frac{1}{n-1}\sum_i (x_i - \bar{x})^2
$$

```{r}
var(vti)
```

## Spread: using `sd()` for sample standard deviation

$$
sample \; standard \; deviation = s = \sqrt{\frac{1}{n-1}\sum_i (x_i - \bar{x})^2}
$$

```{r}
sd(vti)
```

## Spread: using `scale()` for z-scores

$$
\text{z-scores} = \frac{x_i - \bar{x}}{s}
$$

```{r}
z_scores <- scale(vti)[,1]
```

## Spread: counting outlier proportion from z-scores

```{r}
outliers <- abs(z_scores) > 2
outlier_proportion <- sum(outliers) / length(z_scores)
outlier_proportion
```

## Shape: measuring sample skewness

$$
sample \; skewness = \frac{1}{n}\sum_i z_i^3
$$

## Shape: there is no built-in function for skewness in R

```{r}
skew <- function(x) {
  n <- length(x)
  z_scores <- scale(x)[,1]
  return(sum(z_scores^3) / n)
}
skew(vti)
```

## Shape: measuring sample excess kurtosis

$$
sample \; excess \; kurtosis = \frac{1}{n}\sum_i z_i^4 - 3
$$
## Shape: there is no built-in function for excess kurtosis in R

```{r}
kurtosis <- function(x) {
  n <- length(x)
  z_scores <- scale(x)[,1]
  return(sum(z_scores^4) / n - 3)
}
kurtosis(vti)
```

## Summaries of a categorical vector

- Tabulating by frequencies
- Tabulating by probabilities

## Converting numeric vector to categorical

```{r}
price_diff <- diff(vti)
is_rise <- price_diff >= 0
```

## Using `table()` function to tabulate data

```{r}
table(is_rise)
prop.table(table(is_rise))
```

# Bivariate data

## What is bivariate data

A bivariate data set is a set of measurements for some variable from a collection of cases involving two attributes.

$$
x_1, x_2, ..., x_n \\
y_1, y_2, ..., y_n
$$

## How do we use R for bivariate data

Specifying 2 vectors.

## Getting 2 vectors with real-world data

```{r message=FALSE}
start_date <- "2018-01-01"
getSymbols(Symbols = "AAPL", from = start_date) # create a AAPL xts object in global environment
getSymbols(Symbols = "AMZN", from = start_date) # create a AMZN xts object in global environment
aapl <- as.vector(coredata(AAPL$AAPL.Adjusted))
amzn <- as.vector(coredata(AMZN$AMZN.Adjusted))
```

## Data structures for 2 vectors

- matrix
- list
- data.frame

## Using matrix for 2 vectors

```{r}
aapl_amzn <- cbind(matrix(aapl), matrix(amzn))
class(aapl_amzn)
dim(aapl_amzn)
```

## Using list for 2 vectors

```{r}
aapl_amzn <- list(
  aapl,
  amzn
)
class(aapl_amzn)
```

## Using data.frame for 2 vectors

```{r}
aapl_amzn <- data.frame(
  aapl,
  amzn
)
class(aapl_amzn)
dim(aapl_amzn)
```

## Summaries of 2 numeric vectors

- Covariance
- Correlation

## Using `cov()` for covariance

$$
cov(x, y) = \frac{1}{n-1}\sum(x_i - \bar{x})(y_i - \bar{y})
$$

```{r}
cov(aapl, amzn)
```

## Using `cor()` for correlation

$$
cor(x, y) = \frac{cov(x, y)}{s_x s_y}
$$

```{r}
cor(aapl, amzn)
```

## Summaries of 2 categorical vectors

```{r}
price_diff_aapl <- diff(aapl)
price_diff_amzn <- diff(amzn)
is_rise_aapl <- price_diff_aapl >= 0
is_rise_amzn <- price_diff_amzn >= 0
```

## Using `table()` function to create frequency table

```{r}
freq_table <- table(is_rise_aapl, is_rise_amzn)
freq_table
```

## Using `prop.table()` function to convert frequency table to a proportional table

```{r}
prop_table <- prop.table(freq_table)
prop_table
```

## Using `addmargins()` function for summations

```{r}
addmargins(freq_table)
addmargins(prop_table)
```

# Multivariate data

## What is multivariate data

A multivariate data set is a set of measurements for some variable from a collection of cases involving multiple attributes.

## It is pretty much the same as in bivariate data

## Just notice if the data has heterogeneous or homogeneous features

- Use matrix for homogeneous features
- Use list or data.frame for heterogeneous features

## Use matrix for a homogeneous multivariate data

```{r}
amzn_matrix <- cbind(
  matrix(AMZN$AMZN.Open),
  matrix(AMZN$AMZN.High),
  matrix(AMZN$AMZN.Low),
  matrix(AMZN$AMZN.Close),
  matrix(AMZN$AMZN.Adjusted),
  matrix(AMZN$AMZN.Volume)
)
class(amzn_matrix)
dim(amzn_matrix)
```

## Use data.frame for a heterogeneous multivariate data

```{r}
amzn_df <- data.frame(AMZN)
```

---

```{r}
class(amzn_df)
dim(amzn_df)
str(amzn_df)
```

---

```{r}
aapl_df <- data.frame(AAPL)
```

---

```{r}
class(aapl_df)
dim(aapl_df)
str(aapl_df)
```

## Combining 2 data.frame objects

```{r}
col_names <- c("Open", "High", "Low", "Close", "Volume", "Adjusted")
trading_dates <- row.names(aapl_df)
colnames(aapl_df) <- col_names
colnames(amzn_df) <- col_names
aapl_df$Ticker <- "AAPL"
amzn_df$Ticker <- "AMZN"
aapl_df$Ticker <- as.character(aapl_df$Ticker)
amzn_df$Ticker <- as.character(amzn_df$Ticker)
aapl_df$TradingDate <- trading_dates
amzn_df$TradingDate <- trading_dates
row.names(aapl_df) <- NULL
row.names(amzn_df) <- NULL
```

---

```{r}
combined_df <- rbind(aapl_df, amzn_df)
combined_df <- combined_df[c("TradingDate", "Ticker", "Open", "High", "Low", "Close", "Volume", "Adjusted")]
head(combined_df)
tail(combined_df)
```

---

```{r}
str(combined_df)
```