---
title: "Data in Statistics"
author: 郭耀仁 Yao-Jen Kuo
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    center: true
    reveal_options:
      self_contained: true
      previewLinks: true
      slideNumber: true
---

# What is statistical analysis

## Commons of statistical analysis

- The practice and study of collecting and analyzing data
- A fact about or summary of some data

## Types of statistical analysis

- Descriptive statistics: Describe and summarize (population) data
- Inferential statistics: Use a sample of data to make inferencesabout a larger population

## Types of data

- Numerical data
- Categorical data

## Dimensions of data

- Univariate data
- Bivariate data
- Multivariate data

## Real world data: AAPL

```{r message=FALSE}
library(quantmod)

start_date <- "2018-01-01"
getSymbols(Symbols = "AAPL", from = start_date) # create a AAPL xts object in global environment
```

# Numerical data

## Numerical data: continuous(measured)

```{r}
aapl <- as.vector(coredata(AAPL$AAPL.Adjusted)) # extract adjusted price as a vector
class(aapl)
```

## Numerical data: discrete(counted)

The rises and falls of AAPL.

```{r}
aapl_is_rise <- as.numeric(diff(aapl) >= 0)
class(aapl_is_rise)
aapl_is_rise[1:5]
```

## Categorical data: nominal

```{r}
tickers <- c("AAPL", "AMZN", "TSLA")
class(tickers)
```

## Categorical data: ordered

```{r}
suggestions_char <- c("Strong sell", "Sell", "Hold", "Buy", "Strong buy")
suggestions <- factor(suggestions_char, ordered = TRUE, levels = suggestions_char)
class(suggestions)
suggestions
as.numeric(suggestions)
```

# Univariate data

## How do we use R for univariate data

Vector.

$$
x_1, x_2, ..., x_n
$$

## Summaries of a numeric vector

- Measures of center
- Measures of spread
- Measures of shape

## Measures of center

- `mean()`
- `median()`
- `quantile()`

## `mean()` function

$$
\bar{x} = \frac{1}{n}(x_1 + x_2 + ... + x_n) = \frac{1}{n}\sum_i x_i
$$

---

```{r}
my_mean <- function(x) {
  n <- length(x)
  summation <- sum(x)
  return(summation / n)
}
```

---

```{r}
my_mean(aapl)
mean(aapl)
```

---

## `median()` function

The middle value of the sorted vector.

```{r}
my_median <- function(x) {
  n <- length(x)
  sorted_x <- sort(x)
  middle_index <- n %/% 2 + 1 # floor divide
  if (n %% 2 == 1) { # modulo
    return(sorted_x[middle_index])
  } else {
    middle_indices <- c(middle_index - 1, middle_index)
    middle_values <- sorted_x[middle_indices]
    return(sum(middle_values) / 2)
  }
}
```

---

```{r}
my_median(aapl)
median(aapl)
```

## `quantile()` function

A more generalized version of `median()` function.

```{r}
quantile(aapl)
```

## Measures of spread

- `range()` function
- `IQR()` function
- `var()` function
- `sd()` function
- `scale()` function

## `range()` function

The difference between the smallest and largest values.

```{r}
max(aapl) - min(aapl)
diff(range(aapl))
```

## `IQR()` function

The difference between 75% quantile and 25% quantile, which prevents being affected by just one large or small value.

```{r}
aapl_quantile <- quantile(aapl)
unname(aapl_quantile["75%"] - aapl_quantile["25%"])
IQR(aapl)
```

## Outlier

Data point that is substantially different from the others.

$$
data < Q1−1.5 \times IQR \\
data > Q3+1.5 \times IQR
$$

## Finding outliers

```{r}
lower_threshold <- quantile(aapl, 0.25) - 1.5 * IQR(aapl)
upper_threshold <- quantile(aapl, 0.75) + 1.5 * IQR(aapl)
unname(lower_threshold)
unname(upper_threshold)
aapl[aapl < lower_threshold | aapl > upper_threshold]
```

## `var()` function

Average distance from each data point to the data's mean.

$$
s^2 = \frac{1}{n-1}\sum_i (x_i - \bar{x})^2
$$

---

```{r}
my_var <- function(x) {
  n <- length(x)
  x_bar <- mean(x)
  squared_error <- (x - x_bar)**2
  sum_squared_error <- sum(squared_error)
  variance <- sum_squared_error / (n - 1)
  return(variance)
}
```

---

```{r}
my_var(aapl)
var(aapl)
```

## `sd()`function

$$
s = \sqrt{\frac{1}{n-1}\sum_i (x_i - \bar{x})^2}
$$

---

```{r}
my_sd <- function(x) {
  return(sqrt(my_var(x)))
}
```

---

```{r}
my_sd(aapl)
sd(aapl)
```

## `scale()` function

$$
\text{z-scores} = \frac{x_i - \bar{x}}{s}
$$

---

```{r}
my_scale <- function(x) {
  x_bar <- mean(x)
  s <- sd(x)
  z_scores <- (x - x_bar) / s
  return(z_scores)
}
```

---

```{r}
my_scale(aapl)[1:5]
scale(aapl)[1:5]
```

## Measures of shape

- Skewness
- Kurtosis

## Skewness measures the degree of symmetry

- Positive skewness means a right-tailed shape
- Negative skewness means a left-tailed shape

$$
Skewness = \frac{1}{n}\sum_i z_i^3
$$

---

```{r}
skewness <- function(x) {
  n <- length(x)
  z_scores <- scale(x)[,1]
  return(sum(z_scores^3) / n)
}
skewness(aapl)
```

## Kurtosis measures the amount of potential outliers

$$
Kurtosis = \frac{1}{n}\sum_i z_i^4 - 3
$$

---

```{r}
kurtosis <- function(x) {
  n <- length(x)
  z_scores <- scale(x)[,1]
  return(sum(z_scores^4) / n - 3)
}
kurtosis(aapl)
```

## Summaries of a categorical vector

- Tabulating by frequencies
- Tabulating by probabilities

## Converting numeric vector to categorical

```{r}
price_diff <- diff(aapl)
is_rise <- price_diff >= 0
```

## Using `table()` function to tabulate data

```{r}
table(is_rise)
prop.table(table(is_rise))
```

# Bivariate data

## What is bivariate data

A bivariate data set is a set of measurements for some variable from a collection of cases involving two attributes.

$$
x_1, x_2, ..., x_n \\
y_1, y_2, ..., y_n
$$

## How do we use R for bivariate data

Specifying 2 vectors.

## Getting 2 vectors with real-world data

```{r message=FALSE}
start_date <- "2018-01-01"
getSymbols(Symbols = "AAPL", from = start_date) # create a AAPL xts object in global environment
getSymbols(Symbols = "AMZN", from = start_date) # create a AMZN xts object in global environment
aapl <- as.vector(coredata(AAPL$AAPL.Adjusted))
amzn <- as.vector(coredata(AMZN$AMZN.Adjusted))
```

## Data structures for 2 vectors

- matrix
- list
- data.frame

## Using matrix for 2 vectors

```{r}
aapl_amzn <- cbind(matrix(aapl), matrix(amzn))
class(aapl_amzn)
dim(aapl_amzn)
```

## Using list for 2 vectors

```{r}
aapl_amzn <- list(
  aapl,
  amzn
)
class(aapl_amzn)
```

## Using data.frame for 2 vectors

```{r}
aapl_amzn <- data.frame(
  aapl,
  amzn
)
class(aapl_amzn)
dim(aapl_amzn)
```

## Summaries of 2 numeric vectors

- Covariance
- Correlation

## Using `cov()` for covariance

$$
cov(x, y) = \frac{1}{n-1}\sum(x_i - \bar{x})(y_i - \bar{y})
$$

---

```{r}
my_cov <- function(x, y) {
  x_bar <- mean(x)
  y_bar <- mean(y)
  n <- length(x)
  error <- (x - x_bar)*(y-y_bar)
  sum_error <- sum(error)
  covariance <- sum_error / (n-1)
  return(covariance)
}
```

---

```{r}
my_cov(aapl, amzn)
cov(aapl, amzn)
```

## Using `cor()` for correlation

$$
cor(x, y) = \frac{cov(x, y)}{s_x s_y}
$$

---

```{r}
my_cor <- function(x, y) {
  cov_xy <- my_cov(x, y)
  sd_x <- my_sd(x)
  sd_y <- my_sd(y)
  correlation <- cov_xy / (sd_x*sd_y)
  return(correlation)
}
```

---

```{r}
my_cor(aapl, amzn)
cor(aapl, amzn)
```

## Summaries of 2 categorical vectors

```{r}
price_diff_aapl <- diff(aapl)
price_diff_amzn <- diff(amzn)
is_rise_aapl <- price_diff_aapl >= 0
is_rise_amzn <- price_diff_amzn >= 0
```

## Using `table()` function to create frequency table

```{r}
freq_table <- table(is_rise_aapl, is_rise_amzn)
freq_table
```

## Using `prop.table()` function to convert frequency table to a proportional table

```{r}
prop_table <- prop.table(freq_table)
prop_table
```

## Using `addmargins()` function for summations

```{r}
addmargins(freq_table)
addmargins(prop_table)
```

# Multivariate data

## What is multivariate data

A multivariate data set is a set of measurements for some variable from a collection of cases involving multiple attributes.

## Just notice if the data has heterogeneous or homogeneous features

- matrix for homogeneous features
- list or data.frame for heterogeneous features

## Using matrix for a homogeneous multivariate data

```{r}
amzn_matrix <- cbind(
  matrix(AMZN$AMZN.Open),
  matrix(AMZN$AMZN.High),
  matrix(AMZN$AMZN.Low),
  matrix(AMZN$AMZN.Close),
  matrix(AMZN$AMZN.Adjusted),
  matrix(AMZN$AMZN.Volume)
)
class(amzn_matrix)
dim(amzn_matrix)
```

## Using data.frame for a heterogeneous multivariate data

```{r}
amzn_df <- data.frame(AMZN)
```

---

```{r}
class(amzn_df)
dim(amzn_df)
str(amzn_df)
```

---

```{r}
aapl_df <- data.frame(AAPL)
```

---

```{r}
class(aapl_df)
dim(aapl_df)
str(aapl_df)
```

## Combining 2 data.frame objects

```{r}
col_names <- c("Open", "High", "Low", "Close", "Volume", "Adjusted")
trading_dates <- row.names(aapl_df)
colnames(aapl_df) <- col_names
colnames(amzn_df) <- col_names
aapl_df$Ticker <- "AAPL"
amzn_df$Ticker <- "AMZN"
aapl_df$Ticker <- as.character(aapl_df$Ticker)
amzn_df$Ticker <- as.character(amzn_df$Ticker)
aapl_df$TradingDate <- trading_dates
amzn_df$TradingDate <- trading_dates
row.names(aapl_df) <- NULL
row.names(amzn_df) <- NULL
```

---

```{r}
combined_df <- rbind(aapl_df, amzn_df)
combined_df <- combined_df[c("TradingDate", "Ticker", "Open", "High", "Low", "Close", "Volume", "Adjusted")]
head(combined_df)
tail(combined_df)
```

---

```{r}
str(combined_df)
```

## Real world data: Covid 19 daily report

```{r}
read_daily_report <- function(file_date) {
  file_date_str <- format(file_date, "%m-%d-%Y")
  file_url <- sprintf("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/%s.csv", file_date_str)
  daily_report <- read.csv(file_url, stringsAsFactors = FALSE)
  return(daily_report)
}
sys_date <- Sys.Date() - 2
daily_report <- read_daily_report(sys_date)
```

---

```{r}
str(daily_report)
```