---
title: "Exploring Data in Statistics"
author: 郭耀仁 Yao-Jen Kuo
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    center: true
    reveal_options:
      self_contained: true
      previewLinks: true
      slideNumber: true
---

## Agenda

- Exploring univariate data
- Exploring bivariate data
- Exploring multivariate data

# Exploring univariate data

## Getting a vector with real-world data

```{r message=FALSE}
library(quantmod)

start_date <- "2018-01-01"
getSymbols(Symbols = "VTI", from = start_date) # create a VTI xts object in global environment
vti <- as.vector(coredata(VTI$VTI.Close)) # extract close prices as a vector
class(vti)
```

## Using `hist()` to explore the shape of numeric univariate data

```{r}
hist(vti, breaks = 30)
```

## Using density plot to explore the shape of numeric univariate data

```{r}
plot(density(vti))
```

## Combining histogram and density plot to explore the shape of numeric univariate data

```{r}
dens <- density(vti)
hist(vti, probability = TRUE, breaks = 30)
lines(dens)
```

## Using quantile plot to explore the shape of numeric univariate data

```{r}
qqnorm(vti)
```

## Using box plot to explore the shape of numeric univariate data

```{r}
boxplot(vti)
```

## Using bar plot to explore tabulating of categorical univariate data

```{r}
price_diff <- diff(vti)
is_rise <- price_diff >= 0
table_is_rise <- table(is_rise)
```

---

```{r}
barplot(table_is_rise, horiz = TRUE)
```

# Exploring bivariate data

## Getting 2 vectors with real-world data

```{r message=FALSE}
start_date <- "2018-01-01"
getSymbols(Symbols = "AAPL", from = start_date) # create a AAPL xts object in global environment
getSymbols(Symbols = "AMZN", from = start_date) # create a AMZN xts object in global environment
aapl <- as.vector(coredata(AAPL$AAPL.Close))
amzn <- as.vector(coredata(AMZN$AMZN.Close))
```

## Using scatter to explore the relationship of numeric bivariate data

```{r}
plot(aapl, amzn)
```

# Exploring multivariate data