---
title: "Exploring Data in Statistics"
author: 郭耀仁 Yao-Jen Kuo
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    center: true
    reveal_options:
      self_contained: true
      previewLinks: true
      slideNumber: true
---

# Exploring Data in a Nutshell

## The purpose of data exploration

- Exploring shapes
- Exploring correlations
- Exploring ranks
- Exploring trends

## Basic plots for theses explorations

- Exploring shapes via histogram / boxplot / quantile plot
- Exploring correlations via scatter plot
- Exploring ranks via bar plot
- Exploring trends via line plot

## Basic exploration tools in R

- Base plotting system
- ggplot2

```{r}
library(ggplot2)
```

## Getting real-world data: VTI

```{r message=FALSE}
library(quantmod)

start_date <- "2018-01-01"
getSymbols(Symbols = "VTI", from = start_date) # create a VTI xts object in global environment
vti <- as.vector(coredata(VTI$VTI.Adjusted)) # extract adjusted prices as a vector
class(vti)
```

## Getting real-world data: AAPL

```{r message=FALSE}
getSymbols(Symbols = "AAPL", from = start_date) # create a AAPL xts object in global environment
aapl <- as.vector(coredata(AAPL$AAPL.Adjusted)) # extract adjusted prices as a vector
class(aapl)
```

## Getting real-world data: AMZN

```{r message=FALSE}
start_date <- "2018-01-01"
getSymbols(Symbols = "AMZN", from = start_date) # create a AMZN xts object in global environment
amzn <- as.vector(coredata(AMZN$AMZN.Adjusted))
```

## Getting real-world data: Covid 19 daily report

```{r}
read_daily_report <- function(file_date) {
  file_date_str <- format(file_date, "%m-%d-%Y")
  file_url <- sprintf("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports/%s.csv", file_date_str)
  daily_report <- read.csv(file_url, stringsAsFactors = FALSE)
  return(daily_report)
}
sys_date <- Sys.Date() - 2
daily_report <- read_daily_report(sys_date)
```

## Getting real-world data: Covid 19 time series

```{r}
read_time_series <- function() {
  file_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
  time_series <- read.csv(file_url, stringsAsFactors = FALSE)
  return(time_series)
}
time_series <- read_time_series()
``` 

# Exploring univariate data

## Using `hist()` to explore the shape of numeric univariate data

```{r}
hist(vti, breaks = 30)
```

---

```{r}
hist(aapl, breaks = 30)
```

---

```{r}
library(dplyr)

confirmed_by_country <- daily_report %>% 
  group_by(Country_Region) %>% 
  summarise(confirmed = sum(Confirmed)) %>% 
  as.data.frame()
```

---

```{r}
hist(confirmed_by_country$confirmed, breaks = 30)
```

## Using density plot to explore the shape of numeric univariate data

```{r}
plot(density(vti))
```

---

```{r}
plot(density(aapl))
```

---

```{r}
plot(density(confirmed_by_country$confirmed))
```

## Combining histogram and density plot to explore the shape of numeric univariate data

```{r}
dens <- density(vti)
hist(vti, probability = TRUE, breaks = 30)
lines(dens)
```

---

```{r}
dens <- density(aapl)
hist(aapl, probability = TRUE, breaks = 30)
lines(dens)
```

## Using quantile plot to explore the shape of numeric univariate data

```{r}
qqnorm(vti)
```

---

```{r}
qqnorm(aapl)
```

---

```{r}
qqnorm(confirmed_by_country$confirmed)
```

## Using box plot to explore the shape of numeric univariate data

```{r}
boxplot(vti)
```

---

```{r}
boxplot(aapl)
```

## Using bar plot to explore tabulating of categorical univariate data

```{r}
diff_vti <- diff(vti)
is_rise <- diff_vti > 0
table(is_rise)
barplot(table(is_rise), horiz = TRUE, main = "VTI")
```

---

```{r}
diff_aapl <- diff(aapl)
is_rise <- diff_aapl > 0
table(is_rise)
barplot(table(is_rise), horiz = TRUE, main = "AAPL")
```

---

```{r}
n_rows <- table(daily_report[["Country_Region"]])
n_rows <- sort(n_rows, decreasing = TRUE)
barplot(n_rows[2:10], las=2)
```

# Exploring bivariate data

## Using line plot to explore the trend of bivariate data

---

```{r}
plot(aapl)
```

---

```{r}
plot(amzn)
```

## Adjusting the data format of Covid 19 time series

```{r}
library(tidyr)

col_names <- colnames(time_series)
cols_to_pivot <- col_names[5:length(col_names)]
time_series_longer <- time_series %>% 
  pivot_longer(cols = cols_to_pivot, names_to = "Date", values_to = "Confirmed") %>% 
  as.data.frame()
```

## Transforming date column from character to Date

```{r}
date_col <- time_series_longer[["Date"]]
date_col <- sub(pattern = "X", replacement = "", date_col)
date_col <- as.Date(date_col, format = "%m.%d.%y")
time_series_longer[["Date"]] <- date_col
str(time_series_longer)
```

---

```{r}
library(dplyr)

time_series_longer %>% 
  filter(Country.Region == 'Taiwan*') %>% 
  ggplot(aes(x = Date, y = Confirmed)) +
    geom_line()
```

## Using scatter plot to explore the relationship of numeric bivariate data

---

```{r}
plot(aapl, amzn)
```

# Exploring multivariate data

## Combining 2 data.frame objects

```{r}
aapl_df <- data.frame(AAPL)
amzn_df <- data.frame(AMZN)
```

---

```{r}
col_names <- c("Open", "High", "Low", "Close", "Volume", "Adjusted")
trading_dates <- row.names(aapl_df)
colnames(aapl_df) <- col_names
colnames(amzn_df) <- col_names
aapl_df$Ticker <- "AAPL"
amzn_df$Ticker <- "AMZN"
aapl_df$Ticker <- as.character(aapl_df$Ticker)
amzn_df$Ticker <- as.character(amzn_df$Ticker)
aapl_df$TradingDate <- trading_dates
amzn_df$TradingDate <- trading_dates
row.names(aapl_df) <- NULL
row.names(amzn_df) <- NULL
```

---

```{r}
combined_df <- rbind(aapl_df, amzn_df)
combined_df <- combined_df[c("TradingDate", "Ticker", "Open", "High", "Low", "Close", "Volume", "Adjusted")]
```

## Using other aesthetics for multivariate data

```{r}
# different line colors
combined_df$TradingDate <- base::as.Date(combined_df$TradingDate)
p <- ggplot(combined_df, aes(x = TradingDate, y = Adjusted, col = Ticker)) +
  geom_line()
```

---

```{r}
p
```

---

```{r}
# different line colors
p <- time_series_longer %>% 
  filter(Country.Region %in% c("Taiwan*", "Japan", "Korea, South", "Singapore")) %>% 
  ggplot(aes(x = Date, y = Confirmed, col = Country.Region)) +
  geom_line()
```

---

```{r}
p
```

---

```{r}
# different boxes
p <- ggplot(combined_df, aes(x = Ticker, y = Adjusted, fill = Ticker)) + 
  geom_boxplot()
```

---

```{r}
p
```

---

```{r}
# different boxes
p <- ggplot(combined_df, aes(x = Ticker, y = Adjusted, fill = Ticker)) + 
  geom_boxplot()
```

---

```{r}
# different histograms
p <- ggplot(combined_df, aes(x = Adjusted, fill = Ticker)) + 
  geom_histogram(bins = 30) +
  facet_wrap(~ Ticker, scales = "free")
```

---

```{r}
p
```