---
title: "Exploring Data in Statistics"
author: 郭耀仁 Yao-Jen Kuo
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    center: true
    reveal_options:
      self_contained: true
      previewLinks: true
      slideNumber: true
---

# Exploring Data in a Nutshell

## The purpose of data exploration

- Exploring shapes
- Exploring correlations
- Exploring ranks
- Exploring trends

## Basic plots for theses explorations

- Exploring shapes via histogram / boxplot / quantile plot
- Exploring correlations via scatter plot
- Exploring ranks via bar plot
- Exploring trends via line plot

## Basic exploration tools in R

- Base plotting system
- ggplot2

```{r}
library(ggplot2)
```

# Exploring univariate data

## Getting a vector with real-world data

```{r message=FALSE}
library(quantmod)

start_date <- "2018-01-01"
getSymbols(Symbols = "VTI", from = start_date) # create a VTI xts object in global environment
vti <- as.vector(coredata(VTI$VTI.Close)) # extract close prices as a vector
class(vti)
```

## Using `hist()` to explore the shape of numeric univariate data

```{r}
hist(vti, breaks = 30)
```

## Using density plot to explore the shape of numeric univariate data

```{r}
plot(density(vti))
```

## Combining histogram and density plot to explore the shape of numeric univariate data

```{r}
dens <- density(vti)
hist(vti, probability = TRUE, breaks = 30)
lines(dens)
```

## Using quantile plot to explore the shape of numeric univariate data

```{r}
qqnorm(vti)
```

## Using box plot to explore the shape of numeric univariate data

```{r}
boxplot(vti)
```

## Using bar plot to explore tabulating of categorical univariate data

```{r}
price_diff <- diff(vti)
is_rise <- price_diff >= 0
table_is_rise <- table(is_rise)
```

---

```{r}
barplot(table_is_rise, horiz = TRUE)
```

# Exploring bivariate data

## Getting 2 vectors with real-world data

```{r message=FALSE}
start_date <- "2018-01-01"
getSymbols(Symbols = "AAPL", from = start_date) # create a AAPL xts object in global environment
getSymbols(Symbols = "AMZN", from = start_date) # create a AMZN xts object in global environment
aapl <- as.vector(coredata(AAPL$AAPL.Close))
amzn <- as.vector(coredata(AMZN$AMZN.Close))
```

## Using line plot to explore the trend of bivariate data

---

```{r}
plot(AAPL$AAPL.Close)
```

---

```{r}
plot(AMZN$AMZN.Close)
```

## Using scatter plot to explore the relationship of numeric bivariate data

---

```{r}
plot(aapl, amzn)
```

# Exploring multivariate data

## Getting real-world multivariate data

```{r message=FALSE}
start_date <- "2018-01-01"
getSymbols(Symbols = "AAPL", from = start_date) # create a AAPL xts object in global environment
getSymbols(Symbols = "AMZN", from = start_date) # create a AMZN xts object in global environment
aapl_df <- data.frame(AAPL)
amzn_df <- data.frame(AMZN)
```

## Combining 2 data.frame objects

```{r}
col_names <- c("Open", "High", "Low", "Close", "Volume", "Adjusted")
trading_dates <- row.names(aapl_df)
colnames(aapl_df) <- col_names
colnames(amzn_df) <- col_names
aapl_df$Ticker <- "AAPL"
amzn_df$Ticker <- "AMZN"
aapl_df$Ticker <- as.character(aapl_df$Ticker)
amzn_df$Ticker <- as.character(amzn_df$Ticker)
aapl_df$TradingDate <- trading_dates
amzn_df$TradingDate <- trading_dates
row.names(aapl_df) <- NULL
row.names(amzn_df) <- NULL
```

---

```{r}
combined_df <- rbind(aapl_df, amzn_df)
combined_df <- combined_df[c("TradingDate", "Ticker", "Open", "High", "Low", "Close", "Volume", "Adjusted")]
```

## Using other aesthetics for multivariate data

```{r}
# different line colors
combined_df$TradingDate <- as.Date(combined_df$TradingDate)
p <- ggplot(combined_df, aes(x = TradingDate, y = Adjusted, col = Ticker)) +
  geom_line()
```

---

```{r}
p
```

---

```{r}
# different dot colors
p <- ggplot(combined_df, aes(x = Volume, y = Adjusted, col = Ticker)) + 
  geom_point()
```

---

```{r}
p
```

---

```{r}
# different boxes
p <- ggplot(combined_df, aes(x = Ticker, y = Adjusted, col = Ticker)) + 
  geom_boxplot()
```

---

```{r}
p
```