---
title: "Essential R Programming"
author: 郭耀仁 Yao-Jen Kuo
date: "`r Sys.Date()`"
output:
  revealjs::revealjs_presentation:
    theme: night
    highlight: zenburn
    center: true
    reveal_options:
      self_contained: true
      previewLinks: true
      slideNumber: true
---

## Agenda

- Basic Syntax
- Built-in Data Structures
- Intermediate Syntax

# Basic Syntax

## The symbol of assignment `<-`

```{r}
hello_msg <- "Hello from R!"
hello_msg
```

## The symbol of comments `#`

```{r}
# Saying hello from R
hello_msg <- "Hello from R!"
hello_msg # print hello_msg to console
```

## Style guide for R programming

<http://adv-r.had.co.nz/Style.html>

## Useful built-in functions

- `print()`
- `help()`
- `rm()`
- `getwd()`
- `sessionInfo()`

## Common vector types

- numeric
- character
- logical
- Date

## Numeric with length of 1

```{r}
numeric_vector <- 5566
numeric_vector
class(numeric_vector)
length(numeric_vector)
```

## Numeric with length larger than 1

```{r}
numeric_vector <- c(5, 5, 6, 6) # using help(c) or ?c to check what c() function is
numeric_vector
class(numeric_vector)
length(numeric_vector)
```

## Character with length of 1

```{r}
character_vector <- "五五六六"
character_vector
class(character_vector)
length(character_vector)
```

## Character with length larger than 1

```{r}
character_vector <- c("五", "五", "六", "六")
character_vector
class(character_vector)
length(character_vector)
```

## Logical with length of 1

```{r}
logical_vector <- TRUE
logical_vector
class(logical_vector)
length(logical_vector)
```

## Date with length of 1

```{r}
date_vector <- Sys.Date()
date_vector
class(date_vector)
length(date_vector)
```

## Date with length larger than 1

```{r}
date_vector <- c(Sys.Date(), Sys.Date() - 1, Sys.Date() - 2, Sys.Date() - 3)
date_vector
class(date_vector)
length(date_vector)
```

## if statement

```{r eval=FALSE}
if (LOGICAL) {
  # Codes written here will be ran once LOGICAL is evaluated as TRUE
  # ...
  # ...
}
```

---

```{r}
wanna_fight <- TRUE
if (wanna_fight) {
  print("要打，就去練舞室打！")
}
```

## if-else statement

```{r eval=FALSE}
if (LOGICAL) {
  # Codes written here will be ran once LOGICAL is evaluated as TRUE
} else {
  # Codes written here will be ran once LOGICAL is evaluated as FALSE
}
```

---

```{r}
wanna_fight <- FALSE
if (wanna_fight) {
  print("要打，就去練舞室打！")
} else {
  print("那就不要打架。")
}
```

## if-else if-else statement

```{r eval=FALSE}
if (LOGICAL_1) {
  # Codes written here will be ran once LOGICAL_1 is evaluated as TRUE
} else if (LOGICAL_2) {
  # Codes written here will be ran once LOGICAL_1 is evaluated as FALSE
  # And LOGICAL_2 is evaluated as TRUE
} else {
  # Codes written here will be ran once both LOGICAL_1 and LOGICAL_2 are evaluated as TRUE
}
```

---

```{r}
wanna_fight <- FALSE
dancing_room_is_open <- FALSE
if (wanna_fight) {
  print("要打，先看練舞室有沒有開放。")
} else if (dancing_room_is_open) {
  print("要打，練舞室又有開放就去練舞室打！")
} else {
  print("那就不要打架。")
}
```

# Built-in Data Structures

## An overview of built-in data structures

|Number of Dimensions|Homogeneous|Heterogeneous|
|--------------------|-----------|-------------|
|1|Vector|List|
|2|Matrix|Data frame|
|n|Array||

# Intermediate Syntax

## Iterating with `for`

## Iterating with `while`

## Defining functions